---
title: "Final Project"
output:
  pdf_document: default
  html_document: default
---

# Final Project

## Imports

```{r echo=FALSE}
# API
library(httr)
library(jsonlite)

# Time Series
library(forecast)
library(tseries)

# Plotting
library(ggplot2)
library(gridExtra)
```

## API Initialization

```{r}
credentials_file <- "CREDENTIALS.json"
credentials <- fromJSON(credentials_file)

api_key <- credentials$CRYPTOCOMPARE_API_KEY
```

### API Test Query

```{r}
url <- "https://min-api.cryptocompare.com/data/price"
parameters <- list(
  fsym = "BTC",
  tsyms = "USD" 
)

headers <- add_headers(
  authorization = paste("Apikey", api_key)  # Add your API key to the header
)
response <- GET(url, query=parameters, headers)
```

```{r}
data <- content(response, as="text", encoding="UTF-8")
parsed_data <- fromJSON(data, flatten=TRUE)
print(parsed_data)
```

## Question 2

How well can an ARIMA model predict the short-term price movements of altcoins (e.g., Ethereum, Solana, Sui, and Litecoin) based on historical Bitcoin price data and which altcoin follows Bitcoin the most?

### Load Data

```{r}
url <- "https://min-api.cryptocompare.com/data/v2/histohour"
toTs <- as.numeric(as.POSIXct("2024-11-24 00:00:00", tz="UTC"))
limit <- 7 * 24
```

#### Bitcoin (BTC)

```{r}
parameters <- list(
  fsym = "BTC",
  tsym = "USD",
  limit = limit,
  toTs = toTs
)
response <- GET(url, query=parameters, headers)
```

```{r}
data <- content(response, as = "parsed", encoding = "UTF-8")
btc_prices <- data$Data$Data
btc_prices_df <- data.frame(
  time = as.POSIXct(sapply(btc_prices, function(x) x$time), origin="1970-01-01", tz="UTC"),
  high = sapply(btc_prices, function(x) x$high),
  low = sapply(btc_prices, function(x) x$low),
  open = sapply(btc_prices, function(x) x$open),
  close = sapply(btc_prices, function(x) x$close)
)
btc_prices_df
```

#### Ethereum (ETH)

```{r}
parameters <- list(
  fsym = "ETH",
  tsym = "USD",
  limit = limit,
  toTs = toTs
)
response <- GET(url, query=parameters, headers)
```

```{r}
data <- content(response, as = "parsed", encoding = "UTF-8")
eth_prices <- data$Data$Data
eth_prices_df <- data.frame(
  time = as.POSIXct(sapply(eth_prices, function(x) x$time), origin="1970-01-01", tz="UTC"),
  high = sapply(eth_prices, function(x) x$high),
  low = sapply(eth_prices, function(x) x$low),
  open = sapply(eth_prices, function(x) x$open),
  close = sapply(eth_prices, function(x) x$close)
)
eth_prices_df
```

#### Solana (SOL)

```{r}
parameters <- list(
  fsym = "SOL",
  tsym = "USD",
  limit = limit,
  toTs = toTs
)
response <- GET(url, query=parameters, headers)
```

```{r}
data <- content(response, as = "parsed", encoding = "UTF-8")
sol_prices <- data$Data$Data
sol_prices_df <- data.frame(
  time = as.POSIXct(sapply(sol_prices, function(x) x$time), origin="1970-01-01", tz="UTC"),
  high = sapply(sol_prices, function(x) x$high),
  low = sapply(sol_prices, function(x) x$low),
  open = sapply(sol_prices, function(x) x$open),
  close = sapply(sol_prices, function(x) x$close)
)
sol_prices_df
```

#### Sui Network (SUI)

```{r}
parameters <- list(
  fsym = "SUI",
  tsym = "USD",
  limit = limit,
  toTs = toTs
)
response <- GET(url, query=parameters, headers)
```

```{r}
data <- content(response, as = "parsed", encoding = "UTF-8")
sui_prices <- data$Data$Data
sui_prices_df <- data.frame(
  time = as.POSIXct(sapply(sui_prices, function(x) x$time), origin="1970-01-01", tz="UTC"),
  high = sapply(sui_prices, function(x) x$high),
  low = sapply(sui_prices, function(x) x$low),
  open = sapply(sui_prices, function(x) x$open),
  close = sapply(sui_prices, function(x) x$close)
)
sui_prices_df
```

#### Litecoin (LTC)

```{r}
parameters <- list(
  fsym = "LTC",
  tsym = "USD",
  limit = limit,
  toTs = toTs
)
response <- GET(url, query=parameters, headers)
```

```{r}
data <- content(response, as = "parsed", encoding = "UTF-8")
ltc_prices <- data$Data$Data
ltc_prices_df <- data.frame(
  time = as.POSIXct(sapply(ltc_prices, function(x) x$time), origin="1970-01-01", tz="UTC"),
  high = sapply(ltc_prices, function(x) x$high),
  low = sapply(ltc_prices, function(x) x$low),
  open = sapply(ltc_prices, function(x) x$open),
  close = sapply(ltc_prices, function(x) x$close)
)
ltc_prices_df
```

#### Create Data Frame

```{r}
coin_data <- data.frame(
  time = btc_prices_df$time,
  btc_close = btc_prices_df$close,
  eth_close = eth_prices_df$close,
  sol_close = sol_prices_df$close, 
  sui_close = sui_prices_df$close,
  ltc_close = ltc_prices_df$close
)
coin_data
```

### EDA

```{r}
btc_plot <- (ggplot(coin_data) + theme_minimal()
  + geom_smooth(aes(x=time, y=btc_close), color="red")
  + labs(title="BTC Price", x="BTC ($)", y="Time (Hourly)"))

btc_plot
```

```{r}
eth_plot <- (ggplot(coin_data) + theme_minimal()
  + geom_smooth(aes(x=time, y=eth_close), color="purple")
  + labs(title="ETH Price", x="ETH ($)", y="Time (Hourly)"))

eth_plot
```

```{r}
sol_plot <- (ggplot(coin_data) + theme_minimal()
  + geom_smooth(aes(x=time, y=sol_close), color="blue")
  + labs(title="SOL Price", x="SOL ($)", y="Time (Hourly)"))

sol_plot
```

```{r}
sui_plot <- (ggplot(coin_data) + theme_minimal()
  + geom_smooth(aes(x=time, y=sui_close), color="darkgreen")
  + labs(title="SUI Price", x="SUI ($)", y="Time (Hourly)"))

sui_plot
```

```{r}
ltc_plot <- (ggplot(coin_data) + theme_minimal()
  + geom_smooth(aes(x=time, y=ltc_close), color="orange")
  + labs(title="LTC Price", x="LTC ($)", y="Time (Hourly)"))

ltc_plot
```

### ADF Test

```{r}
adf_results <- lapply(coin_data[,-1], adf.test)

adf_summary <- data.frame(
  coin = names(adf_results), 
  adf_statistic = sapply(adf_results, function(x) x$statistic), 
  p_value = sapply(adf_results, function(x) x$p.value),
  stationary = sapply(adf_results, function(x) ifelse(
    x$p.value < 0.-5, "Stationary", "Non-Stationary"))
)

adf_summary
```

### ACF and PACF Plots

#### BTC Plots

```{r}
btc_acf <- acf(coin_data$btc_close, lag=50, main="BTC ACF Plot")
btc_pacf <- pacf(coin_data$btc_close, main="BTC PACF Plot")
```

#### ETH Plots

```{r}
eth_acf <- acf(coin_data$eth_close, lag=50, main="ETH ACF Plot")
eth_pacf <- pacf(coin_data$eth_close, main="ETH PACF Plot")
```

#### SOL Plots

```{r}
sol_acf <- acf(coin_data$sol_close, lag=50, main="SOL ACF Plot")
sol_pacf <- pacf(coin_data$sol_close, main="SOL PACF Plot")
```

#### SUI Plots

```{r}
sui_acf <- acf(coin_data$sui_close, lag=50, main="SUI ACF Plot")
sui_pacf <- pacf(coin_data$sui_close, main="SUI PACF Plot")
```

#### LTC Plots

```{r}
ltc_acf <- acf(coin_data$ltc_close, lag=50, main="LTC ACF Plot")
ltc_pacf <- pacf(coin_data$ltc_close, main="LTC PACF Plot")
```

### Seasonal Decomposition

#### BTC Decomposition

```{r}
btc_ts <- ts(coin_data$btc_close, frequency = 24)
btc_decomp <- decompose(btc_ts, type = "additive")
plot(btc_decomp)
```

#### ETH Decomposition

```{r}
eth_ts <- ts(coin_data$eth_close, frequency = 24)
eth_decomp <- decompose(eth_ts, type = "additive")
plot(eth_decomp)
```

#### SOL Decomposition

```{r}
sol_ts <- ts(coin_data$sol_close, frequency = 24)
sol_decomp <- decompose(sol_ts, type = "additive")
plot(sol_decomp)
```

#### SUI Decomposition

```{r}
sui_ts <- ts(coin_data$sui_close, frequency = 24)
sui_decomp <- decompose(sui_ts, type = "additive")
plot(sui_decomp)
```

#### LTC Decomposition

```{r}
ltc_ts <- ts(coin_data$ltc_close, frequency = 24)
ltc_decomp <- decompose(ltc_ts, type = "additive")
plot(ltc_decomp)
```

### Time Series Models

```{r}
train_size <- length(btc_ts) - (2*24)

btc_train <- btc_ts[1:train_size]
btc_test <- btc_ts[(train_size+1):length(btc_ts)]

forecast_steps <- length(btc_test)
```

#### ETH Model

```{r}
eth_train <- eth_ts[1:train_size]
eth_test <- eth_ts[(train_size + 1):length(eth_ts)]
```

```{r}
eth_arimax <- auto.arima(eth_train, xreg=btc_train, seasonal=TRUE)

summary(eth_arimax)
```

```{r}
eth_forecast <- forecast(eth_arimax, xreg=btc_test, h=forecast_steps)

plot(eth_forecast, 
     main="ETH Close Forecast (SARIMAX with BTC)")
lines((train_size+1):length(eth_ts), eth_test, col="red", lwd=2)
legend("topleft", legend=c("Forecast", "Actual"), col=c("blue","red"), lty=1, lwd=2, cex=0.5)
```

```{r}
eth_acc <- accuracy(eth_forecast$mean, eth_test)
print(eth_acc)
```

#### SOL Model

```{r}
sol_train <- sol_ts[1:train_size]
sol_test <- sol_ts[(train_size + 1):length(sol_ts)]
```

```{r}
sol_arimax <- auto.arima(sol_train, xreg=btc_train, seasonal=TRUE)

summary(sol_arimax)
```

```{r}
sol_forecast <- forecast(sol_arimax, xreg=btc_test, h=forecast_steps)

plot(sol_forecast, 
     main="SOL Close Forecast (SARIMAX with BTC)")
lines((train_size+1):length(sol_ts), sol_test, col="red", lwd=2)
legend("topleft", legend=c("Forecast", "Actual"), col=c("blue","red"), lty=1, lwd=2, cex=0.5)
```

```{r}
sol_acc <- accuracy(sol_forecast$mean, sol_test)
print(sol_acc)
```

#### SUI Model

```{r}
sui_train <- sui_ts[1:train_size]
sui_test <- sui_ts[(train_size + 1):length(sui_ts)]
```

```{r}
sui_arimax <- auto.arima(sui_train, xreg=btc_train, seasonal=TRUE)

summary(sui_arimax)
```

```{r}
sui_forecast <- forecast(sui_arimax, xreg=btc_test, h=forecast_steps)

plot(sui_forecast, 
     main="SUI Close Forecast (SARIMAX with BTC)")
lines((train_size+1):length(sui_ts), sui_test, col="red", lwd=2)
legend("topleft", legend=c("Forecast", "Actual"), col=c("blue","red"), lty=1, lwd=2, cex=0.5)
```

```{r}
sui_acc <- accuracy(sui_forecast$mean, sui_test)
print(sui_acc)
```

#### LTC Model

```{r}
ltc_train <- ltc_ts[1:train_size]
ltc_test <- ltc_ts[(train_size + 1):length(ltc_ts)]
```

```{r}
ltc_arimax <- auto.arima(ltc_train, xreg=btc_train, seasonal=TRUE)

summary(ltc_arimax)
```

```{r}
ltc_forecast <- forecast(ltc_arimax, xreg=btc_test, h=forecast_steps)

plot(ltc_forecast, 
     main="LTC Close Forecast (SARIMAX with BTC)")
lines((train_size+1):length(ltc_ts), ltc_test, col="red", lwd=2)
legend("topleft", legend=c("Forecast", "Actual"), col=c("blue","red"), lty=1, lwd=2, cex=0.5)
```

```{r}
ltc_acc <- accuracy(ltc_forecast$mean, ltc_test)
print(ltc_acc)
```

### Correlation of Altcoins to BTC

```{r}
coin_data$btc_return <- c(NA, diff(coin_data$btc_close) / head(coin_data$btc_close, -1)) * 100
coin_data$eth_return <- c(NA, diff(coin_data$eth_close) / head(coin_data$eth_close, -1)) * 100
coin_data$sol_return <- c(NA, diff(coin_data$sol_close) / head(coin_data$sol_close, -1)) * 100
coin_data$sui_return <- c(NA, diff(coin_data$sui_close) / head(coin_data$sui_close, -1)) * 100
coin_data$ltc_return <- c(NA, diff(coin_data$ltc_close) / head(coin_data$ltc_close, -1)) * 100

```

```{r}
correlation_btc_eth <- cor(coin_data$btc_return, coin_data$eth_return, use = "complete.obs")
correlation_btc_sol <- cor(coin_data$btc_return, coin_data$sol_return, use = "complete.obs")
correlation_btc_sui <- cor(coin_data$btc_return, coin_data$sui_return, use = "complete.obs")
correlation_btc_ltc <- cor(coin_data$btc_return, coin_data$ltc_return, use = "complete.obs")
```

```{r}
cat("Correlation between BTC and ETH:", correlation_btc_eth, "\n")
cat("Correlation between BTC and SOL:", correlation_btc_sol, "\n")
cat("Correlation between BTC and SUI:", correlation_btc_sui, "\n")
cat("Correlation between BTC and LTC:", correlation_btc_ltc, "\n")
```
